### 2019-08-17 thesis.tex
the end?

% Considering the context presented above, the working proposal of
% this PhD Thesis is to put forward the first numerical analysis of
% the new weak turbulence theory for collisional plasmas \cite{YZKS16}
% and explore its applications in the solar physics scope. In
% \Cref{cha:theo-fram} and \Cref{cha:weak-turb}, we make a brief review
% of the basic plasma kinetic theory equations and the weak turbulence
% theory, respectively. In \Cref{noneigenmode}, we introduce the
% new generalization of the weak turbulence theory, which takes into
% account the contribution of noneigenmode fluctuations during the
% derivation of the particle and wave kinetic equations, resulting
% in three new terms. In \Cref{main}, we contextualize the present
% work regarding the theory discussed in the previous chapter and
% present the results obtained so far by appending the two papers
% published in this subject. In \Cref{cha:final}, we summarize the
% results and discuss the perspectives of the present research. Then,
% we propose a schedule for the remaining period until the thesis
% defense. In the end we have three appendices. \Cref{appA} shows
% an improved approximation for the second order susceptibility of
% the electrostatic bremsstrahlung for Langmuir waves. \Cref{appB}
% presents a discussion about the asymptotic equilibrium state of the
% Langmuir waves spectrum, considering a combination of a inverse
% power-law kappa distribution and a Maxwellian distribution, in the
% presence and in the absence of the electrostatic bremsstrahlung.
% The discussion in \Cref{appB} is related with the article appended
% in \Cref{appC}, which is a parallel research and is only slightly
% related with the main subject of this work, that is the study of the
% combined action of collective and collisional processes in weakly
% turbulent plasmas, in the presence of electrostatic oscillations.

\chapter{Kinetic theory of plasmas}
\label{sec:kin-theo}

% The plasma is a collection of ions, electrons and neutral particles that has
% its dynamics dominated by the electric and magnetic fields, generated by the
% electrically charged particles and the electric currents related to their
% motion, respectively. Due to the long-range nature of these self-generated
% fields, one particle affects the movement of several other particles, which
% will interact with other particles, and so on. This \emph{collective behavior}
% \footnote{A detailed exposition of all the processes and conditions that
%   characterize the collective behavior in plasmas, can be accessed in the
%   following introductory books \cite{bitt,gurnett2017,chen,DavidsonNonNeutral}.}
% is what characterizes an ionized gas as a plasma, and is also the reason
% behind the complexity of the plasma dynamics. This complexity comes in the
% form of a myriad of self-consistent oscillatory processes, where the electric
% and magnetic fields affect - and are affected by - the position and motion of
% the charges, and how they vary in time \cite{chen}.

% A complete description of such a complicated dynamics would depend on determining
% the position and momentum of each particle, and their respective microscopic
% electric and magnetic fields, in a given instant in time. This strategy is not
% practical and in most of the cases not feasible for many-body systems, like the
% plasma. However, this microscopic description can be dealt through a hierarchy of
% approximations, leading to the three main theories of plasma physics: kinetic,
% multi-fluid, and magnetohydrodynamics (MHD). The use of each one of these
% formulations will depend on the phenomenon that needs to be described.

The kinetic theory is the most fundamental description of the
plasma dynamics. Based on the concepts and methods of statistical
mechanics, together with the Maxwell equations, it provides a
formal self-consistent evolution of the plasma processes. This
formalism takes into account the variations of the velocity
distribution function $f({\bf r},{\bf v},t)$ in the phase-space
for each particle species. Statistically, this means that the
product between the distribution function and the volume element
in the hexa-dimensional phase-space $f_\alpha({\bf r},{\bf v},t)d^3r d^3v$
represents the probability of finding a particle of species
$\alpha$ in a volume element $d^3r$ around ${\bf r}$ and in a
velocity volume element $d^3v$ around ${\bf v}$, at the instant
$t$ \cite{klimo,klimon}.

@@

\chapter{Applications to astrophysical plasmas}
\label{appD}
% On Earth, the plasma is considered an exotic state of matter, associated to the
% glowing gas inside fluorescent and neon lamps, microchip manufacturing, research
% laboratories, and the promise of cleaner and practically inexhaustible energy
% source once we have controlled and energetically efficient thermonuclear fusion
% \cite{bell}. Naturally occurring plasmas in our everyday life are rare, restricted
% to the occurrence of lightning during storms, flames, and the beautiful auroras
% that take place at high latitudes. In the space, however, the situation is the
% opposite. Starting in the outer layers of Earth's atmosphere, passing through the
% solar wind in the interplanetary space, to the stellar interiors and atmospheres,
% accretion disks, and molecular clouds it is estimated that something around
% $99 \sim 99.9\%$ of the observable matter in the universe is in the plasma state
% \cite{may}. So understanding the basic physics of plasma processes might help on
% interpreting and describing

% The analysis presented in this doctoral thesis has relevant contributions in the
% study of the combined action of nonlinear kinetic instabilities and collisional
% processes in unmagnetized, fully ionized, hydrogen plasmas. Though the plasma's
% profile under consideration might be considered quite idealized, the physical
% processes that are being addressed have fundamental character - the plasma waves
% under consideration and the particle interactions are inherent to any plasma.
% The underlying nature of these processes means that they might impact the plasma
% dynamics at different astrophysical environments and also in laboratory plasmas.
% Laboratory plasmas, however, are usually constrained by specific experimental
% set-up. This specificity would require in-depth knowledge and maybe access to
% one experiment, which makes the analysis in the experimental context restrictive
% and counterproductive. A more broad approach was then to compare our theoretical
% results with phenomena observed in the naturally occurring space plasma.

% \footnote{In the literature, the term ``astrophysical
%   plasma'' is sometimes used to refer to the plasma observed outside the solar
%   system, while ``space plasma'' is used to account for the plasma in our
%   interplanetary space. In this text, however, both terms are being used as
%   synonyms.}

% \footnote{A comprehensive description of the scales and characteristics of
%   cosmic plasmas can be seen in Section 1.2 of Ref. \cite{peratt2014}.}

Astrophysical plasmas are observed in a vast range of densities, temperatures,
magnetic fields, ionization degree, and scale lengths\cite{peratt2014}. Starting
in the outer layers of Earth's atmosphere, passing through the solar wind in the
interplanetary space, to the stellar interiors and atmospheres, accretion disks,
and molecular clouds it is estimated that something around $99 \sim 99.9\%$ of the
observable matter in the universe is in the plasma state. Under cosmic conditions,
even the weakly ionized gas of the neutral hydrogen regions around galaxies or in
the atmosphere of cold stars have a strong reaction to electromagnetic fields,
exhibiting the characteristic collective behavior of plasmas and, therefore, are
also considered plasmas\cite{tsypa,may,peratt2014}.

@@

The theory developed during this graduate research analyses the combined
action of nonlinear kinetic instabilities and collisional processes in
the context of solar physics. To contextualize our work within the solar
plasma research, in this appendix, we present a brief review of the
current state-of-the-art of solar and space physics research, followed
by a discussion on the solar plasma phenomena related to our study.

\section{Solar and space physics}
\label{sec:solar-physics}
The domain of solar and space physics is the \emph{heliosphere}, a cavity
in the interstellar medium, created by the constant supersonic flow of solar
plasma that permeates the interplanetary space. This region encloses the
solar system and extends for approximately $140\unit{AU}$\footnote{An AU
  (astronomical unit) is defined as the mean distance between the Sun and
  Earth, $1\unit{AU}\approx 150\times 10^6\unit{km}$.} in the surrounding
galactic space. Its outer limit is the \emph{heliopause}, an interface
zone where the outgoing pressure of the solar plasma balances the incoming
pressure from the interstellar medium. This comet-shaped plasma bubble
\cite{Dialynas2017}, also has an inner boundary, the \emph{termination shock},
which delimits the sphere of absolute influence of the solar plasma. Beyond
the termination shock, at roughly $100\unit{AU}$ from the Sun, extending
out to the heliopause, lies the \emph{heliosheath}. In this region, the
solar wind particles start being subjected to increasing external pressure,
transitioning to a subsonic, turbulent flow that blends and interacts
directly with the matter in the local interstellar medium \cite{NAP13060}.
In \autoref{figD1}, we have a depiction of the regions and boundaries
described above, and the current location of both Voyager missions, that
are already outside the heliosphere, in the interstellar space.
\begin{figure}[h]
  \begin{center}
    \includegraphics[width=0.892\textwidth]{solar_system_env.png}
    \caption{Main boundaries of the heliosphere and the location of both
      Voyager missions.}
      {\footnotesize Credit:}
      \href{https://voyager.jpl.nasa.gov/news/details.php?article_id=112}
      {\footnotesize NASA/JPL-Caltech}{\footnotesize , accessed at 08/08/2019.}
    \label{figD1}
  \end{center}
\end{figure}

@@

\subsection{The Sun and the near-Earth environment}
\label{sec:sun}
A better comprehension of the Sun and the dynamic space around our planet
goes beyond the pure scientific curiosity. This accessible space physics
laboratory can indeed lead to discoveries that can be applied and extended
to remote cosmic systems. However, the primary reason behind the international
effort\footnote{A complete description, of NASA's missions is available at
\url{https://science.nasa.gov/missions-page?field_division_tid=5&field_phase_tid=All}.
For ESA's missions, go to \url{http://sci.esa.int/home/51459-missions/}.
Both websites were accessed at 11/08/2019.} (see Figures \ref{nasa} and
\ref{esa}) on understanding the complex physics of the Sun and interplanetary
space is to unlock means of predicting the occurrence and intensity of solar
activity, and anticipate how the near-Earth environment will be affected.
Reliable forecasting of the space weather is essential for protecting
electronic equipment, radio communications, GPS signal and humans in space
from the effects of the interaction between the dense, magnetized plasma of
coronal mass ejections with the geomagnetic field, and also from the
energetic radiation emitted by solar flares. Extreme coronal mass ejections
can also cause stronger geomagnetic disturbances that may affect terrestrial
infrastructure such as electric power grids \cite{NAP13060}.
\begin{figure}[hb]
  \begin{center}
    \includegraphics[width=0.892\textwidth]{hso_fleet_chart_april_2019.pdf}
    \caption{NASA's present and future heliophysics system observatory. The number
      between parenthesis under the probe's name is the total number of spacecrafts
      that compose the mission.}
      {\footnotesize Credit:}
      \href{https://science.nasa.gov/heliophysics}
      {\footnotesize NASA}{\footnotesize , accessed at 10/08/2019.}
    \label{nasa}
  \end{center}
\end{figure}
\begin{figure}[ht]
  \begin{center}
    \includegraphics[width=0.892\textwidth]{ESA_s_fleet_of_Solar_System_explorers_2019.jpg}
    \caption{ESA's past, present, and future science missions for studying the solar system.}
      {\footnotesize Credit:}
      \href{https://www.esa.int/spaceinimages/Images/2019/02/ESA_s_fleet_of_Solar_System_explorers}
      {\footnotesize ESA}{\footnotesize , accessed at 10/08/2019.}
    \label{esa}
  \end{center}
\end{figure}

@@

Counter-intuitively, the solar corona is, at the same time, the most tenuous
and the hottest outer layer of the Sun. At the top of the chromosphere, where
the temperature might reach $20,000\,\unit{K}$, in a relatively narrow interface,
the \emph{solar transition region}, the plasma temperature rises abruptly to
$\sim 10^6\,\unit{K}$ \cite{BenzA2002}. It has been almost eighty years since
the extreme coronal temperature came as the only reasonable explanation for the
origin of the spectral lines observed in the solar corona \cite{Hunter1942},
in a classical case of an answer that brings a more troublesome question. 
Since then, several mechanisms have been proposed \cite{Parker1988,Scudder92b,
  Vinas2000,PP14}, not only to comprehend the heating of the solar corona,
but also the acceleration of the solar wind. Another unsolved problem that is
directly connected to the inverted temperature profile of the solar corona, is
regarding where and how the non-Maxwellian velocity distribution functions with
suprathermal tails, observed in the solar wind, are formed.

In \cite{Tigik2017a} ...
present a new fundamental mechanism that is capable
the formation of suprathermal tails in the electron velocity distribution function

proposing a fundamental mechanism,
the electrostatic bremsstrahlung, that might take electrons out of the thermal
equilibrium

\begin{figure}[hb]
  \begin{center}
    \includegraphics[width=0.892\textwidth]{corona2017.png}
    \includegraphics[width=0.892\textwidth]{corona2019.png}
    \caption{Solar corona captured during the total solar eclipse of 2017 (top panel)
      and 2019 (lower panel). The reddish contour at some regions is the chromosphere,
      and the reddish arcs are solar prominences, chromospheric plasma structures that
      penetrate the coronal region.
      {\footnotesize Credit:}
      \href{https://hdr-astrophotography.com}
      {\footnotesize Nicolas Lefaudeux}{\footnotesize , accessed at 15/08/2019}.}
    \label{corona}
  \end{center}
\end{figure}

>>>>

Considering the context presented above, the working proposal of
this PhD Thesis is to put forward the first numerical analysis of
the new weak turbulence theory for collisional plasmas \cite{YZKS16}
and explore its applications in the solar physics scope. In
\Cref{cha:theo-fram} and \Cref{cha:weak-turb}, we make a brief review
of the basic plasma kinetic theory equations and the weak turbulence
theory, respectively. In \Cref{noneigenmode}, we introduce the
new generalization of the weak turbulence theory, which takes into
account the contribution of noneigenmode fluctuations during the
derivation of the particle and wave kinetic equations, resulting
in three new terms. In \Cref{main}, we contextualize the present
work regarding the theory discussed in the previous chapter and
present the results obtained so far by appending the two papers
published in this subject. In \Cref{cha:final}, we summarize the
results and discuss the perspectives of the present research. Then,
we propose a schedule for the remaining period until the thesis
defense. In the end we have three appendices. \Cref{appA} shows
an improved approximation for the second order susceptibility of
the electrostatic bremsstrahlung for Langmuir waves. \Cref{appB}
presents a discussion about the asymptotic equilibrium state of the
Langmuir waves spectrum, considering a combination of a inverse
power-law kappa distribution and a Maxwellian distribution, in the
presence and in the absence of the electrostatic bremsstrahlung.
The discussion in \Cref{appB} is related with the article appended
in \Cref{appC}, which is a parallel research and is only slightly
related with the main subject of this work, that is the study of the
combined action of collective and collisional processes in weakly
turbulent plasmas, in the presence of electrostatic oscillations.
%%% I still have to list the last two appendixes in introduction

\chapter{Kinetic theory of plasmas}
\label{sec:kin-theo}
The kinetic theory is the most fundamental description of the
plasma dynamics. Based on the concepts and methods of statistical
mechanics, together with the Maxwell equations, it provides a
formal self-consistent evolution of the plasma processes. This
formalism takes into account the variations of the velocity
distribution function $f({\bf r},{\bf v},t)$ in the phase-space
for each particle species. Statistically, this means that the
product between the distribution function and the volume element
in the hexa-dimensional phase-space $f_\alpha({\bf r},{\bf v},t)d^3r d^3v$
represents the probability of finding a particle of species
$\alpha$ in a volume element $d^3r$ around ${\bf r}$ and in a
velocity volume element $d^3v$ around ${\bf v}$, at the instant
$t$ \cite{klimo,klimon}.

@@

\chapter{Two-dimensional time evolution of beam-plasma instability in the
  presence of binary collisions}
\label{master}
Results of the Master's dissertation that led to the current doctoral
research.
\includepdf[pages=-]{Tigik2016a.pdf}

\chapter{Applications to astrophysical plasmas}
\label{appD}
Astrophysical plasmas are observed in a vast range of densities, temperatures,
magnetic fields, ionization degree, and scale lengths\cite{peratt2014}. Starting
in the outer layers of Earth's atmosphere, passing through the solar wind in the
interplanetary space, to the stellar interiors and atmospheres, accretion disks,
and molecular clouds it is estimated that something around $99 \sim 99.9\%$ of the
observable matter in the universe is in the plasma state. Under cosmic conditions,
even the weakly ionized gas of the neutral hydrogen regions around galaxies or in
the atmosphere of cold stars have a strong reaction to electromagnetic fields,
exhibiting the characteristic collective behavior of plasmas and, therefore, are
also considered plasmas\cite{tsypa,may,peratt2014}.

@@

The theory developed during this graduate research analyses the combined
action of nonlinear kinetic instabilities and collisional processes in
the context of solar physics. To contextualize our work within the solar
plasma research, in \autoref{sec:solar-physics} we present a brief review
of the current research in solar and space physics and the phenomena
related to this work. The corresponding paper is pointed in a footnote.

\section{Solar and space physics}
\label{sec:solar-physics}
The domain of solar and space physics is the \emph{heliosphere}, a cavity
in the interstellar medium, created by the constant supersonic flow of solar
plasma that permeates the interplanetary space. This region encloses the
solar system and extends for approximately $140\unit{AU}$\footnote{An AU
  (astronomical unit) is defined as the mean distance between the Sun and
  Earth, $1\unit{AU}\approx 150\times 10^6\unit{km}$.} in the surrounding
galactic space. Its outer limit is the \emph{heliopause}, an interface
zone where the outgoing pressure of the solar plasma balances the incoming
pressure from the interstellar medium. This comet-shaped plasma bubble
\cite{Dialynas2017}, also has an inner boundary, the \emph{termination shock},
which delimits the sphere of absolute influence of the solar plasma. Beyond
the termination shock, at roughly $100\unit{AU}$ from the Sun, extending
out to the heliopause, lies the \emph{heliosheath}. In this region, the
solar wind particles start being subjected to increasing external pressure,
transitioning to a subsonic, turbulent flow that blends and interacts
directly with the matter in the local interstellar medium \cite{NAP13060}.
In \autoref{figE1}, we have a depiction of the regions and boundaries
described above, and the current location of both Voyager missions, that
are already outside the heliosphere, in the interstellar space.
\begin{figure}[h]
  \begin{center}
    \includegraphics[width=0.892\textwidth]{solar_system_env.png}
    \caption{Main boundaries of the heliosphere and the location of both
      Voyager missions.}
      {\footnotesize Credit:}
      \href{https://voyager.jpl.nasa.gov/news/details.php?article_id=112}
      {\footnotesize NASA/JPL-Caltech}{\footnotesize , accessed at 08/08/2019.}
    \label{figE1}
  \end{center}
\end{figure}

@@

\subsection{The Sun and the near-Earth environment}
\label{sec:sun}
A better comprehension of the Sun and the dynamic space around our planet
goes beyond the pure scientific curiosity. This accessible space physics
laboratory can indeed lead to discoveries that can be applied and extended
to remote cosmic systems. However, the primary reason behind the international
effort (see Figures \ref{nasa} and
\ref{esa}) on understanding the complex physics of the Sun and interplanetary
space is to unlock means of predicting the occurrence and intensity of solar
activity, and anticipate how the near-Earth environment will be affected.
Reliable forecasting of the space weather is essential for protecting
electronic equipment, radio communications, GPS signal and humans in space
from the effects of the interaction between the dense, magnetized plasma of
coronal mass ejections with the geomagnetic field, and also from the
energetic radiation emitted by solar flares\footnote{The emission of hard X-rays
  and type III radio bursts, both generated during solar flares, are addressed in
  \cite{Tigik2016a} (see \autoref{master}).}. Extreme coronal mass ejections
can also cause stronger geomagnetic disturbances that may affect terrestrial
infrastructure such as electric power grids \cite{NAP13060}.
\begin{figure}[hb]
  \begin{center}
    \includegraphics[width=0.892\textwidth]{hso_fleet_chart_april_2019.pdf}
    \caption{NASA's present and future heliophysics system observatory. The number
      between parenthesis under the probe's name is the total number of spacecrafts
      that compose the mission.}
      {\footnotesize Credit:}
      \href{https://science.nasa.gov/heliophysics}
      {\footnotesize NASA}{\footnotesize , accessed at 10/08/2019.}
    \label{nasa}
  \end{center}
\end{figure}
\begin{figure}[ht]
  \begin{center}
    \includegraphics[width=0.892\textwidth]{ESA_s_fleet_of_Solar_System_explorers_2019.jpg}
    \caption{ESA's past, present, and future science missions for studying the solar system.}
      {\footnotesize Credit:}
      \href{https://www.esa.int/spaceinimages/Images/2019/02/ESA_s_fleet_of_Solar_System_explorers}
      {\footnotesize ESA}{\footnotesize , accessed at 10/08/2019.}
    \label{esa}
  \end{center}
\end{figure}

@@

Counter-intuitively, the solar corona is, at the same time, the most tenuous
and the hottest outer layer of the Sun. At the top of the chromosphere, where
the temperature might reach $20,000\,\unit{K}$, in a relatively narrow interface,
the \emph{solar transition region}, the plasma temperature rises abruptly to
$\sim 10^6\,\unit{K}$ \cite{BenzA2002}. It has been almost eighty years since
the extreme coronal temperature came as the only reasonable explanation for the
origin of the spectral lines observed in the solar corona \cite{Hunter1942},
in a classical case of an answer that brings a more troublesome question. 
Since then, several mechanisms have been proposed \cite{Parker1988,Scudder92b,
  Vinas2000,PP14}, not only to comprehend the heating of the solar corona,
but also the acceleration of the solar wind. Another unsolved problem that is
directly connected to the inverted temperature profile of the solar corona and
particle acceleration in the Sun, is regarding where and how the non-Maxwellian
velocity distribution functions with suprathermal tails, observed in the solar
wind, are formed\footnote{The generation of suprathermal electrons in the lower
  corona and its consequences to the coronal heating problem are tackled in
  \cite{Tigik2017a} (see \autoref{sec:gen-supr}).}.
\begin{figure}
  \begin{center}
    \includegraphics[width=0.892\textwidth]{corona2017.png}
    \includegraphics[width=0.892\textwidth]{corona2019.png}
    \caption{Solar corona captured during the total solar eclipse of 2017 (top panel)
      and 2019 (lower panel). The reddish contour at some regions is the chromosphere,
      and the reddish arcs are solar prominences, chromospheric plasma structures that
      penetrate the coronal region.
      {\footnotesize Credit:}
      \href{https://hdr-astrophotography.com}
      {\footnotesize Nicolas Lefaudeux}{\footnotesize , accessed at 15/08/2019}.}
    \label{corona}
  \end{center}
\end{figure}

%% Eu acho que ainda tenho que escrever algo aqui. Vou pensar ainda a respeito.
